---
title: "Assignment 1"
author: "Alicia Chimeno & Silvia Ferrer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clear space and Load dataset

```{r}
# Libraries
library(tidyverse)
library(skimr)
library(FactoMineR)
library(car)
library(lmtest)
library(ISLR)
library(arm)
library(mice)
library(corrplot)
library(dplyr)

rm(list=ls())
par(mfrow=c(1,1))

#setwd("C:/Users/Silvia/OneDrive - Universitat Politècnica de Catalunya/Escritorio/UPC/MASTER DS/1A/SIM/SIM_assignment_1")
setwd("/Users/ali/Desktop/MASTER/SIM/PROJECT 1")

train<-read.delim("train.csv", sep=',') 
```

## Explore dataset and Preproces dataset


```{r}
dim(train)
names(train)
str(train)
#skim(train)

## recode incorrect NA's 
train$Alley[which(is.na(train$Alley))] <- "No alley access"
train$BsmtQual[which(is.na(train$BsmtQual))] <- "No Basement"
train$BsmtCond[which(is.na(train$BsmtCond))] <- "No Basement"
train$BsmtExposure[which(is.na(train$BsmtExposure))] <- "No Basement"
train$BsmtFinType1[which(is.na(train$BsmtFinType1))] <- "No Basement"
train$BsmtFinType2[which(is.na(train$BsmtFinType2))] <- "No Basement"
train$FireplaceQu[which(is.na(train$FireplaceQu))] <- "No Fireplace"
train$GarageType[which(is.na(train$GarageType))] <- "No Garage"
train$GarageFinish[which(is.na(train$GarageFinish))] <- "No Garage"
train$GarageCond[which(is.na(train$GarageCond))] <- "No Garage"
train$PoolQC[which(is.na(train$PoolQC))] <- "No Pool"
train$Fence[which(is.na(train$Fence))] <- "No Fence"
train$MiscFeature[which(is.na(train$MiscFeature))] <- "None"
```

# Recode variables to factor

```{r}
# Recode: character to factor
char_cols <- which(sapply(train, is.character))
train[, char_cols] <- lapply(train[, char_cols], as.factor) 

# Recode: numeric to factor
train$MSSubClass <- factor(train$MSSubClass)
train$MoSold <- factor(train$MoSold) # month 
#train$GarageYrBlt<-factor(train$GarageYrBlt) 
#train$OverallCond <- factor(train$OverallCond)
#train$OverallQual <- factor(train$OverallQual)

```

## Target variable exploration "SalePrice"


```{r}
res.con <- condes(train,81)
res.con$quali
shapiro.test(train$SalePrice)  ## Normality test of the target variable  ->>>  # We test the null hypothesis H0: variable is normal, H1: variable not normal.  # The p-value is less than 0.05, therefore we reject H0.  # For that we will not consider the target variable as normally distributed
qqnorm(train$SalePrice)
qqline(train$SalePrice)
hist(train$SalePrice, prob = TRUE, col = 'lightblue', main = 'SalePrice Distribution', xlab = 'SalePrice')
#lines(density(train$SalePrice), col = 'red', lwd = 2)
```

# Selection of 10 factor variables and numerical variables 
We make the following assumptions : 
Choose the 10 categorical variables that are more correlated with the target variable

```{r}
vect<-row.names(res.con$quali)
vect[1:10] 
factor_train<-train[,c(vect[1:10])] # select the first 10 factor variables that are more related with the target
num_cols <- names(train)[sapply(train, is.numeric)]
train2 <- train[,c(num_cols, vect[1:10])]
dim(train2)

```

## Variable analysis
# Analysis functions

```{r}
## numeric variable description function
numeric_description <- function(variable, n_breaks) {
  cat("Summary:\n")
  print(summary(variable))
  
  cat("\nCount of missing values:",sum(is.na(variable)),"\n")
  column_name <- sub(".+\\$", "", deparse(substitute(variable)))
  hist(variable, breaks = n_breaks, freq = F, main = paste("Histograma de", column_name), xlab = column_name)
  curve(dnorm(x, mean(variable), sd(variable)), add = T)
  
  # Normality test with Shapiro-Wilk
  print(shapiro.test(variable))
}

## outliers functions
analyze_outliers <- function(data, column_name) {
  sm <- summary(data[[column_name]])
  iqr <- sm["3rd Qu."] - sm["1st Qu."]
  
  # Mild Outliers
  mild_ub <- sm["3rd Qu."] + 1.5 * iqr
  mild_lb <- sm["1st Qu."] - 1.5 * iqr
  
  mild_outliers <- length(which(data[[column_name]] > mild_ub | data[[column_name]] < mild_lb))
  cat("Number of mild outliers:", mild_outliers, "\n")
  
  # Plotting mild outliers
  Boxplot(data[[column_name]], main = paste("Outlier Analysis for", column_name),
          ylab = column_name, outline = TRUE)
  abline(h = mild_ub, col = "orange", lwd = 2)
  abline(h = mild_lb, col = "orange", lwd = 2)
  
  # Severe Outliers
  severe_ub <- sm["3rd Qu."] + 3 * iqr
  severe_lb <- sm["1st Qu."] - 3 * iqr
  severe_outliers <- length(which(data[[column_name]] > severe_ub | data[[column_name]] < severe_lb))
  cat("Number of severe outliers:", severe_outliers, "\n")
  # Plotting severe outliers
  abline(h = severe_ub, col = "red", lwd = 2.5)
  abline(h = severe_lb, col = "red", lwd = 2.5)
}
# count outliers (new column)
train2$univ_outl_count <- 0 # initialize
update_outliers_count <- function(dataframe, column_name) {
  df <- dataframe
  sm <- summary(df[[column_name]])
  iqr <- sm["3rd Qu."] - sm["1st Qu."]
  # Severe Outliers
  severe_ub <- sm["3rd Qu."] + 3 * iqr
  severe_lb <- sm["1st Qu."] - 3 * iqr
  severe_outliers_id <- which(df[[column_name]] > severe_ub | df[[column_name]] < severe_lb)
  df$univ_outl_count[severe_outliers_id] <- df$univ_outl_count[severe_outliers_id] + 1
  return(df)
}

```

# EDA and univariate outliers of each variable

```{r}
# EDA EXPLORATORY DATA ANALYSIS 
skim(train2) # overall exploration
names(train2)

### Numerical variables
## Id (not considered for the analysis)

## "LotFrontage" Linear feet of street connected to property (cont)
numeric_description(train2$LotFrontage, 30)
# not normal distribution
analyze_outliers(train2, "LotFrontage")
train2 <- update_outliers_count(train2, "LotFrontage")

## "LotArea" Lot size in square feet (cont)
numeric_description(train2$LotArea, 20)
# not normal distribution
analyze_outliers(train2, "LotArea")
train2 <- update_outliers_count(train2, "LotArea")

## "OverallQual" (cont)
numeric_description(train2$OverallQual, 10)
# not normal distribution
analyze_outliers(train2, "OverallQual")
train2 <- update_outliers_count(train2, "OverallQual")

## "YearBuilt" (int)
summary(train2$YearBuilt)
barplot(table(train2$YearBuilt), main = "Distribution of Year Built",xlab = "Year",col = "skyblue")
sum(is.na(train2$YearBuilt))
analyze_outliers(train2, "YearBuilt")
train2 <- update_outliers_count(train2, "YearBuilt")

## "YearRemodAdd" (int) - Remodel date (same as construction date if no remodeling or additions)
summary(train2$YearRemodAdd)
barplot(table(train2$YearRemodAdd), main = "Distribution of Year Remodelized",xlab = "Year",col = "skyblue")
sum(is.na(train2$YearRemodAdd))
analyze_outliers(train2, "YearRemodAdd")
train2 <- update_outliers_count(train2, "YearRemodAdd")


## MasVnrArea - Masonry veneer area in square feet (cont)
numeric_description(train2$MasVnrArea, 10)
# not normal distribution
analyze_outliers(train2,"MasVnrArea")
train2 <- update_outliers_count(train2, "MasVnrArea")
# comparar amb variable MasVnrType
# si el type != None i es 0 -> convertir en NA
train2$MasVnrArea[which(train2$MasVnrArea == 0)] <- NA
none_idx <- which(train$MasVnrType == 'None')
train2$MasVnrArea[none_idx] <- 0
#Boxplot(train2$MasVnrArea)
#sm<-summary(train2$MasVnrArea)
#iqr <- sm["3rd Qu."]-sm["1st Qu."]; iqr
#train2$f.MasVnrArea <- ifelse(train2$MasVnrArea <= sm["1st Qu."], 1, ifelse(train2$MasVnrArea > sm["1st Qu."] & train2$MasVnrArea < sm["Median"], 2, ifelse(train2$MasVnrArea >= sm["Median"] & train2$MasVnrArea <= sm["3rd Qu."], 3, ifelse(train2$MasVnrArea > sm["3rd Qu."], 4,0))))
#train2$f.MasVnrArea <- factor(train2$f.MasVnrArea,labels=c("LowmasVnrArea","LowMidmasVnrArea","HighMidmasVnrArea","HighmasVnrArea"), order = T, levels=c(1,2,3,4))
#table(train2$f.MasVnrArea)


## "BsmtFinSF1" - Type 1 finished square feet (cont)
numeric_description(train2$BsmtFinSF1, 10)
# not normal distribution
analyze_outliers(train2,"BsmtFinSF1")
train2 <- update_outliers_count(train2, "BsmtFinSF1")


# BsmtFinSF2 - Rating of basement finished area (if multiple types) (cont)
numeric_description(train2$BsmtFinSF2, 10)
# not normal distribution
analyze_outliers(train2,"BsmtFinSF2")
train2 <- update_outliers_count(train2, "BsmtFinSF2")

# "BsmtUnfSF" (cont)
numeric_description(train2$BsmtUnfSF, 20)
# not normal distribution
analyze_outliers(train2,"BsmtUnfSF")
train2 <- update_outliers_count(train2, "BsmtUnfSF")

# "TotalBsmtSF" (cont)
numeric_description(train2$TotalBsmtSF, 10)
# not normal distribution
analyze_outliers(train2,"TotalBsmtSF")
train2 <- update_outliers_count(train2, "TotalBsmtSF")


# X1stFlrSF - First Floor square feet (cont)
numeric_description(train2$X1stFlrSF, 10)
# not normal distribution
analyze_outliers(train2,"X1stFlrSF")
train2 <- update_outliers_count(train2, "X1stFlrSF")

## X2ndFlrSF - Second floor square feet (cont)
numeric_description(train2$X2ndFlrSF, 10)
# not normal distribution
analyze_outliers(train2,"X2ndFlrSF")
train2 <- update_outliers_count(train2, "X2ndFlrSF")
# create a new variable if has second floor = 1 / not second floor = 0. 
train2$secondfloor<-0
train2$secondfloor[which(train2$X2ndFlrSF!=0)]<-1
train2$secondfloor<-factor(train2$secondfloor)
table(train2$secondfloor) 


## LowQualFinSF - Low quality finished square feet (all floors) (cont)
numeric_description(train2$LowQualFinSF, 10)
# not normal distribution
table(train2$LowQualFinSF)# very centered in 0 
analyze_outliers(train2,"LowQualFinSF")
train2 <- update_outliers_count(train2, "LowQualFinSF")

## GrLivArea - Above grade (ground) living area square feet (cont)
numeric_description(train2$GrLivArea, 20)
# not normal distribution
analyze_outliers(train2,"GrLivArea")
train2 <- update_outliers_count(train2, "GrLivArea")

## BsmtFullBath - # full baths in the basement (int)
summary(train2$BsmtFullBath) 
table(train2$BsmtFullBath)
sum(is.na(train2$BsmtFullBath))
barplot(table(train2$BsmtFullBath), main = "Distribution of Baths in the basement",xlab = "Number of Baths",col = "skyblue")
#train2$BsmtFullBath<- factor(train2$BsmtFullBath,labels=c("0","1","2","3 or more"), order = T, levels=c(0,1,2,3))
#plot(train2$BsmtFullBath)

## BsmtHalfBath - # half baths in the basement (int)
summary(train2$BsmtHalfBath) 
table(train2$BsmtHalfBath) # very centered in 0
sum(is.na(train2$BsmtHalfBath))
barplot(table(train2$BsmtHalfBath), main = "Distribution of Half Baths in the basement",xlab = "Number of Baths",col = "skyblue")
#train2$BsmtHalfBath<- factor(train2$BsmtHalfBath,labels=c("0","1","2 or more"), order = T, levels=c(0,1,2))
#plot(train2$BsmtHalfBath)

## FullBath - full baths above grade (int)
summary(train2$FullBath) 
table(train2$FullBath)
sum(is.na(train2$FullBath))
barplot(table(train2$FullBath), main = "Distribution of Baths above grade",xlab = "Number of Baths",col = "skyblue")
#train2$FullBath<- factor(train2$FullBath,labels=c("0","1","2","3 or more"), order = T, levels=c(0,1,2,3))
#plot(train2$FullBath)

## HalfBath - Half baths above grade (int)
summary(train2$HalfBath) 
table(train2$HalfBath)
sum(is.na(train2$HalfBath))
barplot(table(train2$HalfBath), main = "Distribution of Half Baths above grade",xlab = "Number of Baths",col = "skyblue")
#train2$HalfBath<- factor(train2$HalfBath,labels=c("0","1","2 or more"), order = T, levels=c(0,1,2))
#plot(train2$HalfBath)

## BedroomAbvGr - Bedrooms above grade (does NOT include basement bedrooms) (factor?)
summary(train2$BedroomAbvGr) 
table(train2$BedroomAbvGr)
sum(is.na(train2$BedroomAbvGr))
barplot(table(train2$BedroomAbvGr), main = "Distribution of Bedrooms above grade",xlab = "Number of Bedrooms",col = "skyblue")
#train2$BedroomAbvGr<- factor(train2$BedroomAbvGr,labels=c("0","1","2","3","4","5 or more"), order = T, levels=c(0,1,2,3,4,5))
#plot(train2$BedroomAbvGr)

## KitchenAbvGr - Kitchens above grade (int)
summary(train2$KitchenAbvGr) 
table(train2$KitchenAbvGr)
sum(is.na(train2$KitchenAbvGr))
barplot(table(train2$KitchenAbvGr), main = "Distribution of Kitchens",xlab = "Number of Kitchens",col = "skyblue")
#train2$KitchenAbvGr<- factor(train2$KitchenAbvGr,labels=c("0","1","2","3 or more"), order = T, levels=c(0,1,2,3))
#plot(train2$KitchenAbvGr) # very centered variable 

####---------


### SILVIA's half
## "TotRmsAbvGrd" - Total rooms above grade (does not include bathrooms) - continuous ratio variable
summary(train2$TotRmsAbvGrd)
table(train2$TotRmsAbvGrd)
sum(is.na(train2$TotRmsAbvGrd)) #0
Boxplot(train2$TotRmsAbvGrd) # nose
length(Boxplot(train2$TotRmsAbvGrd, id = list(n=Inf))) #30
sevout_TotRmsAbvGrd = (quantile(train2$TotRmsAbvGrd,0.25)+(3*((quantile(train2$TotRmsAbvGrd,0.75)-quantile(train2$TotRmsAbvGrd,0.25)))))
length(which(train2$TotRmsAbvGrd > sevout_TotRmsAbvGrd))
#train2$f.TotRmsAbvGrd <- ifelse(df$mileage <= 5728, 1, ifelse(df$mileage > 5728 & df$mileage <= 16395, 2, ifelse(df$mileage > 16395 & df$mileage <= 33102, 3, ifelse(df$mileage > 33102, 4,0))))
#df$f.mileage <- factor(df$f.mileage,labels=c("LowMileage","LowMidMileage","HighMidMileage","HighMileage"), order = T, levels=c(1,2,3,4))
#table(df$f.mileage)
analyze_outliers(train2,"TotRmsAbvGrd")


## "Fireplaces" - Number of fireplaces (int)
summary(train2$Fireplaces)
table(train2$Fireplaces)
sum(is.na(train2$Fireplaces)) #0
analyze_outliers(train2, "Fireplaces") 
#train2$Fireplaces<- factor(train2$Fireplaces,labels=c("0","1","2","3 or more"), order = T, levels=c(0,1,2,3))
barplot(table(train2$Fireplaces), main = "Distribution of fireplaces",xlab = "Number of Fireplaces",col = "skyblue")

## "GarageCars" - Size of garage in car capacity (int)
summary(train2$GarageCars)
table(train2$GarageCars)
sum(is.na(train2$GarageCars)) #0
#train2$GarageCars <- factor(train2$GarageCars,labels=c("0","1","2","3","4 or more"), order = T, levels=c(0,1,2,3,4))
analyze_outliers(train2, "GarageCars") 
barplot(table(train2$GarageCars), main = "Distribution of Garage Cars",xlab = "Number of Garage Cars",col = "skyblue")


## "GarageArea" - Size of garage in square feet (cont)
numeric_description(train2$GarageArea, 30)
# not normal distribution
analyze_outliers(train2, "GarageArea")
train2 <- update_outliers_count(train2, "GarageArea")
sm<-summary(train2$GarageArea)
train2$f.GarageArea <- ifelse(train2$GarageArea <= sm["1st Qu."], 1, ifelse(train2$GarageArea > sm["1st Qu."] & train2$GarageArea < sm["Median"], 2, ifelse(train2$GarageArea >= sm["Median"] & train2$GarageArea <= sm["3rd Qu."], 3, ifelse(train2$GarageArea > sm["3rd Qu."], 4,0))))
train2$f.GarageArea <- factor(train2$f.GarageArea,labels=c("LowgarageArea","LowMidgarageArea","HighMidgarageArea","HighgarageArea"), order = T, levels=c(1,2,3,4))
table(train2$f.GarageArea)
#checking si los 0 significan que no tienen garage
none_cars <- which(train$GarageCars == 0);none_cars
none_area <- which(train$GarageArea == 0);none_area
none_finish <- which(train$GarageFinish=='No Garage');none_finish

## "WoodDeckSF" - Wood deck area in square feet - continuous ratio variable
numeric_description(train2$WoodDeckSF, 30)
# not normal distribution
table(train2$WoodDeckSF) # centered data in 0
analyze_outliers(train2, "WoodDeckSF")
train2 <- update_outliers_count(train2, "WoodDeckSF")
sm<-summary(train2$WoodDeckSF)
train2$f.WoodDeckSF <- ifelse(train2$WoodDeckSF <= sm["1st Qu."], 1, ifelse(train2$WoodDeckSF > sm["1st Qu."] & train2$WoodDeckSF < sm["Median"], 2, ifelse(train2$WoodDeckSF >= sm["Median"] & train2$WoodDeckSF <= sm["3rd Qu."], 3, ifelse(train2$WoodDeckSF > sm["3rd Qu."], 4,0))))
train2$f.WoodDeckSF <- factor(train2$f.WoodDeckSF,labels=c("LowwoodDeckSF","LowMidwoodDeckSF","HighMidwoodDeckSF","HighwoodDeckSF"), order = T, levels=c(1,2,3,4))
table(train2$f.WoodDeckSF)
# LowMidwoodDeckSF 0 ???

## "OpenPorchSF" - Open porch area in square feet - continuous ratio variable (continua)
#numeric_description(train2$OpenPorchSF, 30)
#analyze_outliers(train2,"OpenPorchSF")
#train2 <- update_outliers_count(train2, "OpenPorchSF")
summary(train2$OpenPorchSF)
table(train2$OpenPorchSF) # very centered data in 0
sum(is.na(train2$OpenPorchSF))
barplot(table(train2$OpenPorchSF), main = "Distribution of ",xlab = "Number of ",col = "skyblue")
sm<-summary(train2$OpenPorchSF)
## faria millor un factor amb tenen porch si/no o bé incluir al d'abaix un q sigui no tenen porch. 
train2$f.OpenPorchSF <- ifelse(train2$OpenPorchSF <= sm["1st Qu."], 1, ifelse(train2$OpenPorchSF > sm["1st Qu."] & train2$OpenPorchSF < sm["Median"], 2, ifelse(train2$OpenPorchSF >= sm["Median"] & train2$OpenPorchSF <= sm["3rd Qu."], 3, ifelse(train2$OpenPorchSF > sm["3rd Qu."], 4,0))))
train2$f.OpenPorchSF <- factor(train2$f.OpenPorchSF,labels=c("LowopenPorchSF","LowMidopenPorchSF","HighMidopenPorchSFF","HighopenPorchSF"), order = T, levels=c(1,2,3,4))
table(train2$f.OpenPorchSF)

## EnclosedPorch - Enclosed porch area in square feet (cont)
summary(train2$EnclosedPorch)
table(train2$EnclosedPorch) # very centered data in 0
sum(is.na(train2$EnclosedPorch))
hist(train2$EnclosedPorch, breaks = 30, freq = F)
# therefore we will create a factor 0 if they dont have enclosed porch area,1 if they do have.
train2$EnclosedPorch_binary<-0
train2$EnclosedPorch_binary[which(train2$EnclosedPorch!=0)]<-1
train2$EnclosedPorch_binary
table(train2$EnclosedPorch_binary)

## X3SsnPorch - Three season porch area in square feet (cont)
numeric_description(train2$X3SsnPorch, 30)
table(train2$X3SsnPorch) # very centered in 0 -> dont have three season porch
analyze_outliers(train2,"X3SsnPorch")
train2 <- update_outliers_count(train2, "X3SsnPorch")


## ScreenPorch - Screen porch area in square feet (cont)
numeric_description(train2$ScreenPorch, 30)
# not normally distributed
table(train2$ScreenPorch) # very centered in 0 -> dont have screen porch
analyze_outliers(train2,"ScreenPorch")
train2 <- update_outliers_count(train2, "ScreenPorch")

## PoolArea - Pool area in square feet (cont)
numeric_description(train2$PoolArea, 30)
# not normally distributed
table(train2$PoolArea) # very centered in 0 -> dont have pool
analyze_outliers(train2,"PoolArea")
train2 <- update_outliers_count(train2, "PoolArea")
# create a new variable : 0=no pool/1= pool
train2$Pool_binary<-0
train2$Pool_binary[which(train2$PoolArea!=0)]<-1
train2$Pool_binary
table(train2$Pool_binary) # not very usefull , the majority dont have pools

## MiscVal - $Value of miscellaneous feature (cont)
numeric_description(train2$MiscVal, 30)
# not normally distributed
table(train2$MiscVal) # very centered in 0 
analyze_outliers(train2,"MiscVal")
train2 <- update_outliers_count(train2, "MiscVal")

## YrSold - (int)
numeric_description(train2$YrSold, 30)
# not normally distributed
table(train2$YrSold) 
analyze_outliers(train2,"YrSold")
train2 <- update_outliers_count(train2, "YrSold")

### categorical variables

## Neighborhood 
sum(is.na(train2$Neighborhood)) #0
table(train2$Neighborhood)
barplot(table(train2$Neighborhood), main = "Distribution of Neighborhood",xlab = "Number of Neighborhood",col = "skyblue")
#pie(table(train2$Neighborhood), main = "Distribution of Neighborhood", col = rainbow(length(levels(train2$Neighborhood))))

## ExterQual - Evaluates the quality of the material on the exterior 
sum(is.na(train2$ExterQual)) #0
table(train2$ExterQual)
barplot(table(train2$ExterQual), main = "Distribution of the exterior material quality",xlab = "Qualification",col = "skyblue")

## BsmtQual - Evaluates the height of the basement
sum(is.na(train2$BsmtQual)) #0
table(train2$BsmtQual)
barplot(table(train2$BsmtQual), main = "Distribution of Basement height Quality",xlab = "Qualification",col = "skyblue")
#pie(table(train2$BsmtQual), main = "Distribution of BsmtQual", col = rainbow(length(levels(train2$BsmtQual))))

## KitchenQual - Kitchen quality
sum(is.na(train2$KitchenQual)) #0
table(train2$KitchenQual)
barplot(table(train2$KitchenQual), main = "Distribution of Kitchen Quality",xlab = "Qualification",col = "skyblue")

## GarageFinish - Interior finish of the garage
sum(is.na(train2$GarageFinish)) #0
table(train2$GarageFinish)
barplot(table(train2$GarageFinish), main = "Garage Finish Distribution",xlab = "Type",col = "skyblue")

## FireplaceQu - Fireplace quality
sum(is.na(train2$FireplaceQu)) #0
table(train2$FireplaceQu)
barplot(table(train2$FireplaceQu), main = "Fireplace quality Distribution",xlab = "Qualification",col = "skyblue")

## Foundation - Type of foundation
sum(is.na(train2$Foundation)) #0
table(train2$Foundation)
barplot(table(train2$Foundation), main = "Type of foundation Distribution",xlab = "Type",col = "skyblue")

## GarageType - Garage location
sum(is.na(train2$GarageType)) #0
table(train2$GarageType)
barplot(table(train2$GarageType), main = "Garage Location Distribution",xlab = "Garage Location",col = "skyblue")

## MSSubClass
sum(is.na(train2$MSSubClass)) #0
table(train2$MSSubClass) # +classes
#barplot(table(train2$MSSubClass), main = "Garage Location Distribution",xlab = "Garage Location",col = "skyblue")
```

# Missing values and imputation

```{r}
## NA's
skim(train2)
sum(is.na(train2$LotFrontage))
sum(is.na(train2$MasVnrArea))
sum(is.na(train2$GarageYrBlt))

## impute
# exclude GarageYrBlt feature because it has no sense to impute this variable

mice_imp<-mice(train2[, !names(train2) %in% "GarageYrBlt"],method = "cart")

#densityplot(train2)
imputed_data<-complete(mice_imp)
# LotFrontage validation
summary(imputed_data$LotFrontage)
summary(train2$LotFrontage)
plot(density(train2$LotFrontage,na.rm=TRUE))
plot(density(imputed_data$LotFrontage,na.rm=TRUE))
# imputation doesnt change much the density nor summary

# MasVnrArea validation
summary(imputed_data$MasVnrArea)
summary(train2$MasVnrArea)
plot(density(train2$MasVnrArea,na.rm=TRUE))
plot(density(imputed_data$MasVnrArea,na.rm=TRUE))
# imputation doesnt change much the density nor summary
```

## Data Quality Exploration

# Correlation Matrix
```{r}
# ORDRE QUE SEGUEIX A L'EXEMPLE
max(imputed_data$univ_outl_count) #9
imputed_data[which(imputed_data$univ_outl_count == 9),]
#df_of_interest = imputed_data[,c(2,3,5,7,8,9,18)] -- canviar nums columnes
#cor_outl = cor(df_of_interest)
#require(corrplot)
#par(mfrow=c(1,1))
#corrplot(cor_outl, method = 'number')

raredata<-imputed_data[which(train2$univ_outl_count>2),]
num_cols <- num_cols[num_cols != "GarageYrBlt"]
cor_outl <- cor(imputed_data[,c(num_cols[-1])])

```


# Multivariate outliers


```{r}
library(chemometrics)
num_cols <- num_cols[num_cols != "Id"]

#Some Moutliers:
# 1. first condes quanti variables
res.con <- condes(imputed_data,35)
vect<-row.names(res.con$quanti)
vect[1:10] 

#cols <- c("EnclosedPorch_binary", "EnclosedPorch_binary")
#n_cols <- num_cols[! num_cols %in% c("EnclosedPorch_binary", "EnclosedPorch_binary") ]
res.out <- Moutlier(imputed_data[,c(vect[1:3])], quantile = 0.9995, col="green")

cor_matrix <- cor(imputed_data[,c(vect[1:3])])
highly_correlated_pairs <- which(cor_matrix > 0.8 & cor_matrix < 1, arr.ind = TRUE)

corrplot(cor_matrix, method = "color")


# 2. Generally the houses built latetly are the ones with most quality
res.out <- Moutlier(imputed_data[,c("YearBuilt","OverallQual")], quantile = 0.9995, col="green")
which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff))
par(mfrow=c(1,1))
plot( res.out$md, res.out$rd )
abline(h=res.out$cutoff, col="red")
abline(v=res.out$cutoff, col="red")

# 3. Generally the houses with more area are the ones with more rooms
res.out <- Moutlier(imputed_data[,c("GrLivArea","TotRmsAbvGrd")], quantile = 0.9995, col="green")
which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff))
par(mfrow=c(1,1))
plot( res.out$md, res.out$rd )
abline(h=res.out$cutoff, col="red")
abline(v=res.out$cutoff, col="red")

# 4. Generally the houses with more total area are the ones with more garage area also
res.out <- Moutlier(imputed_data[,c("GarageArea","TotalBsmtSF")], quantile = 0.9995, col="green")
which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff))
par(mfrow=c(1,1))
plot( res.out$md, res.out$rd )
abline(h=res.out$cutoff, col="red")
abline(v=res.out$cutoff, col="red")

# 5. Moutliers about total areas zones
res.out <- Moutlier(imputed_data[,c("X1stFlrSF","TotalBsmtSF","GrLivArea")], quantile = 0.9995, col="green") #GarageArea, si se lo sumo hay menos outliers
which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff))
par(mfrow=c(1,1))
plot( res.out$md, res.out$rd )
abline(h=res.out$cutoff, col="red")
abline(v=res.out$cutoff, col="red")

# take off m outliers
#df = df[-which((res.out$md > res.out$cutoff)&(res.out$rd > res.out$cutoff)),]

```

# Profiling
To test for autocorrelation the acf() function is used, of which the result can be seen 
below: target is not normally distributed and autocorrelation is found, so a 
randomized dataframe is reset

```{r}
acf(imputed_data$SalePrice)

#duplicate of the section target exploration
shapiro.test(imputed_data$SalePrice)  ## Normality test of the target variable  ->>>  # We test the null hypothesis H0: variable is normal, H1: variable not normal.  # The p-value is less than 0.05, therefore we reject H0.  # For that we will not consider the target variable as normally distributed
qqnorm(imputed_data$SalePrice) + qqline(imputed_data$SalePrice)
hist(imputed_data$SalePrice, prob = TRUE, col = 'lightblue', main = 'SalePrice Distribution', xlab = 'SalePrice') + lines(density(imputed_data$SalePrice), col = 'red', lwd = 2)

ggplot(data=imputed_data, aes(SalePrice, y = after_stat(density))) +
  geom_histogram(breaks=seq(0, max(imputed_data$SalePrice), by=5000),
                col='lightblue',
                fill='steelblue') +
  geom_density(lwd=1,
                col='red') +
  labs(title="Histogram for price with density", x="SalePrice", y="Count")

```
# Variables with most association with the response variable

For the quantitative variables there are some variable that are highly significant positively correlated (r > 0.50, p ~ 0) to the target variable, SalesPrice:  OverallQual, GrLivArea, GarageCars, GarageArea, TotalBsmtSF, X1stFlrSF, FullBath, TotRmsAbvGrd, YearBuilt and YearRemodAdd. This seems logical as the higher the overall condition rating of the house is and the bigger the area above grade living, the more it would cost. There isn't any variable which is highly significant negatively correlated to the SalePrice (r < -0.50 p = 0).

For the qualitative variables it is clear that the Neighborhood explains the most variance in the SalePrice variable (R^2 = 0.546, p ~ 0).  
The influence of the other qualitative variables (with R^2 > 0.3) are in order: Neighborhood, ExterQual, BsmtQual, KitchenQual, f.GarageArea, GarageFinish. Secondfloor is poorly associated as it has a R^2-value under 2%.
-- Then, this factor we have created, we can exclude it.

```{r}
require(FactoMineR)
## Loading required package: FactoMineR
require(corrplot)
res.con <- condes(imputed_data,35) #names(imputed_data)[35] - [1] "SalePrice"
res.con$quanti

#done before for selecting
res.con$quali


```

## Define a polytomic factor ExterQual, evaluates the quality of the material on the exterior, and argue if the average price depends on the level of ExterQual

Define a polytomic factor f.age for the covariate car age according to its quartiles and argue if the average price depends on the level of age. Statistically justify the answer

Given the evidence below, we argue that the price is dependent on the level of ExterQual. Firstly, in the boxplots it can be observed that the better the quality (Q1) the lower the average of the house sale price. This is also seen clearly in the distribution plot below. 
---Secondly, the wilcoxon test shows that the means of these levels are not equal meaning that some relation exists between the factor and the response variable

```{r}
ggplot(imputed_data, aes(x=SalePrice, fill=ExterQual)) +
 geom_density(alpha=.5)

ggplot(imputed_data, aes(y=SalePrice, fill=ExterQual)) +
 geom_boxplot(alpha=0.5)


```

```{r}
pairwise.wilcox.test(imputed_data$SalePrice, imputed_data$ExterQual)

## LowAge LowMidAge HighMidAge
## LowMidAge <2e-16 - - 
## HighMidAge <2e-16 <2e-16 - 
## HighAge <2e-16 <2e-16 <2e-16

#We perform pairwise wilcoxon test to check for similar means. Looking at the boxplot 
#and distribution plot we can already expect that these are not going to be similar. The 
#result of the wilcoxon test indicates our hypothesis was right that there is a clear 
#difference between the means of the different quartiles
```


# Modelling


```{r}
# RSS function
rss <- function(fitted, actual){
  sum((fitted - actual)^2)
}
## Variables selection
# example comparison between models

# variable selection all together
mod.fow <- stats::step(lm(SalePrice ~ ., data = imputed_data), trace = FALSE,
                       direction = "forward")
summary(mod.fow)

display(m0 <- lm(SalePrice ~ 1, data = imputed_data))
display(m1 <- lm(SalePrice ~OverallQual , data = imputed_data))
summary(m0)
rss(fitted(m0), imputed_data$SalePrice)
as.matrix(AIC(m0, m1))
anova(m0,m1) # per veure quin model és millor

## selection of the variables: MSSubClass,
#LotArea, OverallQual, OverallCond, MasVnrArea, BsmtFinSF2
## X1stFlrSF, X2stFlrSF, FullBath, Fireplaces, GarageCars, BsmtQualGd, 

```
